Draft genomes of the three northern hemisphere blue mussel lineages: North and South European \emph{Mytilus edulis} and Mediterranean \emph{Mytilus galloprovincialis}.

### 2022-06-20 manuscript/ms_genomes.tex
barcode filt

We preprocessed 10X reads using the following pipeline for use in several algorithms (\cref{draft_asm}).
We first removed duplicate reads using \texttt{Nubeam-dedup} (\cite{Dai2020}; commit 25dd385).
We then used \texttt{proc10xG} (\url{https://github.com/ucdavis-bioinformatics/proc10xG}; commit 7afbfcf) to split the reads from their 10X barcodes for further processing.
We used a custom filtering step designed to remove under- and over-represented barcodes and their associated reads from the data (\texttt{filter\_barcodes.R} script, see \cite{Bary\&Gagnaire??}).
[[Add supfigs of filter\_barcodes]].
Reads were additionally filtered with \texttt{fastp} (v0.20.1; \cite{Chen2018}) with the main objective to remove poly-G tails created by the Illumina Novaseq sequencing technology.
We obtained at this point reads equivalent to a short-read sequencing run, usable in some parts of the assembly and quality control pipeline.
Additionally, to obtain reads compatible to 10X genomics tools, we filtered and reassembled reads with their 10X barcodes in \texttt{proc10xG} using the \texttt{filter\_10xReads.py} and \texttt{regen\_10xReads.py} scripts.

@@

Preprocessed 10X reads (\cref{preproc}) were used to first estimate estimate genome size and heterozygosities of the three individuals.
We used the reference free k-mer based method \texttt{genomescope} (\url{https://github.com/tbenavi1/genomescope2.0}; commit 5034ed4; \cite{Ranallo-Benavidez2020}) and the fork of the \texttt{KMC} k-mer  counting program (\url{https://github.com/tbenavi1/KMC}; commit 1df71f6).

@@

The pre-assembly k-mer analysis carried out using Genomescope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from 3.49 to 8.77~\% (\cref{fig:genomescope}).
Genomescope provide a bad fit and estimation in the case of the MeduEUS data due to a reduced coverage of the genome in that case, compared to the two other assemblies.

@@

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/Fig1_genomescope.pdf}
	\caption{Genomescope}
	\label{fig:genomescope}
\end{figure}

@@

@@

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/supfig_busco.pdf}
	\caption{Busco scores on the metazoa\_odb10 (left panels) and mollusca\_odb10 (right panels) databases for each assembly MgalMED, MeduEUS and MeduEUN (top to bottom panels) across several versions (v1, v4, v5, v6, v7).}
	\label{supfig:busco}
\end{figure}

>>>>

We preprocessed 10X reads using the following pipeline for use in several algorithms (\cref{draft_asm}).
We first removed duplicate reads using \texttt{Nubeam-dedup} (\cite{Dai2020}; commit 25dd385).
We then used \texttt{proc10xG} (\url{https://github.com/ucdavis-bioinformatics/proc10xG}; commit 7afbfcf) to split the reads from their 10X barcodes for further processing.
We used a custom filtering step designed to remove under- and over-represented barcodes and their associated reads from the data (\cref{supfig:preproc_MgalMED,supfig:preproc_MeduEUS,supfig:preproc_MeduEUN}, \texttt{filter\_barcodes.R} script, see \cite{Bary\&Gagnaire??}).
[[Add supfigs of filter\_barcodes]].
Reads were additionally filtered with \texttt{fastp} (v0.20.1; \cite{Chen2018}) with the main objective to remove poly-G tails created by the Illumina Novaseq sequencing technology.
We obtained at this point reads equivalent to a short-read sequencing run, usable in some parts of the assembly and quality control pipeline.
Additionally, to obtain reads compatible to 10X genomics tools, we filtered and reassembled reads with their 10X barcodes in \texttt{proc10xG} using the \texttt{filter\_10xReads.py} and \texttt{regen\_10xReads.py} scripts.

@@

Preprocessed 10X reads (\cref{preproc}) were used to first estimate estimate genome size and heterozygosities of the three individuals.
We used the reference free k-mer based method \texttt{GenomeScope} (\url{https://github.com/tbenavi1/genomescope2.0}; commit 5034ed4; \cite{Ranallo-Benavidez2020}) and the fork of the \texttt{KMC} k-mer  counting program (\url{https://github.com/tbenavi1/KMC}; commit 1df71f6).

@@

The pre-assembly k-mer analysis carried out using GenomeScope (using 21-mers) showed that, as expected, the genomes were highly heterozygous with values going from 3.49 to 8.77~\% (\cref{fig:genomescope}).
GenomeScope provide a bad fit and estimation in the case of the MeduEUS data due to a reduced coverage of the genome in that case, compared to the two other assemblies.

@@

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/Fig1_genomescope.pdf}
	\caption{k-mer profile plots computed with 21-mers using GenomeScope.
		Coverage histogram of the k-mers in blue.
		Lines represent the fit of the GenomeScope models.
		len: inferred genome length; uniq: percentage of the genome that is unique, aa: overall homozigosity; ab: overall heterozygosity; kcov: mean k-mer coverage for heterozygous bases; err: reads error rate; dup: average rate of read duplication; k: k-mer size; p: ploidy.
	}
	\label{fig:genomescope}
\end{figure}

@@

\subsection*{Tables}

@@

\subsection*{Figures}
% preproc barcodes
\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/MgalMED_preproc_barcode_filt.pdf}
	\caption{Histogram of read pairs identified for each 10X barcode for MgalMED.
		As part of preprocessing step, barcodes for which too few or too many read pairs are associated with each unique barcode are removed from the dataset (red regions).}
	\label{supfig:preproc_MgalMED}
\end{figure}

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/MeduEUS_preproc_barcode_filt.pdf}
	\caption{Histogram of read pairs identified for each 10X barcode for MeduEUS.
		As part of preprocessing step, barcodes for which too few or too many read pairs are associated with each unique barcode are removed from the dataset (red regions).}
	\label{supfig:preproc_MeduEUS}
\end{figure}

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/MeduEUN_preproc_barcode_filt.pdf}
	\caption{Histogram of read pairs identified for each 10X barcode for MeduEUN.
	As part of preprocessing step, barcodes for which too few or too many read pairs are associated with each unique barcode are removed from the dataset (red regions).}
	\label{supfig:preproc_MeduEUN}
\end{figure}

\begin{figure}[h]
	\includegraphics[width=\linewidth]{figures/supfig_busco.pdf}
	\caption{Busco scores on the metazoa\_odb10 (left panels) and mollusca\_odb10 (right panels) databases for each assembly MgalMED, MeduEUS and MeduEUN (top to bottom panels) across several assembly versions (v1, v4, v5, v6, v7).}
	\label{supfig:busco}
\end{figure}
